---
title: "financial analysis"
author: "Mingwei Wu"
output: html_notebook
---


# Data Description
## first_coloumn: Company represents sample companies
## second_column: Time shows different time periods that data belongs to. Time series length varies between 1 to 14 for each company
## third_column: Financial Distress is the target variable. if it is greater than 0.5 the company should be considered as healthy(0). otherwise, it would be regareded as financially distressed (1)
## rest_column: the features denoted by x1 to x83. are some financial and non-financial characteristics of the sampled companies.

# Goal
## checking the dataframe which features are most indicative of financial distress?
## what type of machine learning models perform best of this dataset.



```{r}
library(DataExplorer)
library(caret)
library(knitr)
library(tidyverse)
library(dplyr)
library(ggcorrplot)
library(ggplot2) ## load the package in R
```

```{r}
data<-read.csv("Financial Distress.csv",header = TRUE)
head(data)
```

```{r}
data<-data%>%
  rename(
  company = Company,
  time = Time,
  financial_distress=Financial.Distress
)
```
```{r}
names(data)
```
```{r}
head(data)
```
```{r}
data1<-data[complete.cases(data),]
dim(data1)
```

##plot the relationship betweeen time and financial_distress
```{r}
data1%>%
  ggplot(aes(time,financial_distress))+geom_point(color="blue")
```
```{r}
ggcorrplot(cor(data1))
```

```{r}
plot_histogram(data1)
```
```{r}
summary(data1[1:3])
```
```{r}
summary(data1[4:86])
```
```{r}
pairs(data1[1:3])
```

```{r}
boxplot(data1[4:86]) # most outliers 
```
```{r}
data1<-data1%>%
  select(company,time, financial_distress)%>%
  mutate(distress = ifelse(financial_distress > 0.5 ,1,0))
data1
```













