---
title: "Financial Analysis"
author: "Mingwei Wu"
output: rmarkdown::github_document
---


# Data Description
## first_coloumn: Company represents sample companies
## second_column: Time shows different time periods that data belongs to. Time series length varies between 1 to 14 for each company
## third_column: Financial Distress is the target variable. if it is greater than 0.5 the company should be considered as healthy(0). otherwise, it would be regarded as financially distressed (1)
## rest_column: the features denoted by x1 to x83. are some financial and non-financial characteristics of the sampled companies.

# Goal
## checking the dataframe which features are most indicative of financial distress?
## what type of machine learning models perform best of this dataset.



```{r}
library(DataExplorer)
library(caret)
library(knitr)
library(tidyverse)
library(dplyr)
library(ggcorrplot)
library(gmodels)
library(ROSE)
library(pROC)
library(plyr)
library(randomForest)
library(caTools)
library(ggplot2) ## load the package in R
```

```{r}
data<-read.csv("Financial Distress.csv",header = TRUE)
head(data)
```
```{r}
names(data)
```

```{r}
names(data)[3] <- "financial_distress"
```

```{r}
head(data)
```
```{r}
data1<-data[complete.cases(data),]
dim(data1)
```

### plot the relationship betweeen time and financial_distress
```{r}
data1%>%
  ggplot(aes(Time,financial_distress))+geom_point(color="blue")
```
```{r}
ggcorrplot(cor(data1))
```

```{r}
plot_histogram(data1)
```
```{r}
summary(data1[1:3])
```
```{r}
summary(data1[4:86])
```
```{r}
pairs(data1[1:3])
```

```{r}
boxplot(data1[4:86]) # most outliers 
```
```{r}
data1<-data1%>%
  mutate(distress = ifelse(financial_distress > -0.5 ,0,1))
data1%>%
  select(Company,Time,financial_distress,distress)%>%
  head()
```
```{r}
data1$distress<-as.factor(data1$distress)
```

```{r}
data1%>%
  ggplot(aes(distress))+geom_bar(color="blue",
                                 fill="red")
```

```{r}
glm<-glm(financial_distress~.,data=data1)
summary(glm)  #build up the logistic regression, and most column x is insignificant
```
```{r}
glm1<-glm(financial_distress~Company+Time,data=data1)
summary(glm1) #removed the insignificant columns then make logistic regression
```
```{r}
glm1$coefficients
```

```{r}
set.seed(123)
sample<-createDataPartition(data1$distress,
                            p=0.7,
                            list=FALSE)
train<-data1[sample,]
test<-data1[-sample,] #split data into train and test by 70% and 30%
```
```{r}
dim(train);dim(test)
```
```{r}
balance<-ovun.sample(distress~.,data=train,
                     p=0.5,seed=1,
                     method="both")$data
summary(balance$distress)
```


```{r}
rf1<-randomForest(as.factor(train$distress)~.,train,ntree=100)
rf1
```

```{r}
pred<-predict(rf1, test) #predictions
summary(pred)
```
```{r}
confusionMatrix(pred,as.factor(test$distress),positive="1") #matrix table
```
```{r}
roc<-predict(rf1,test,type="prob")
roc_all<-roc(test$distress, roc[,2])
auc<-as.numeric(round(roc_all$auc,4))
plot(roc_all,lwd=2,col="blue")
text(0.6,0.7,"AUC=")
text(0.4,0.7,auc)
```

















